\subsection{Additional Run Time
  Benchmarks}\label{cockpit::app:run-time-benchmarks}

\subsubsection{Individual Instrument
  Overhead}\label{cockpit::app:benchmark-instruments}
To estimate the computational overhead for individual instruments, we run
\cockpit with that instrument for $32$ iterations, tracking at every step.
Training proceeds with the default batch size specified by the \deepobs problem
and uses \sgd with learning rate $10^{-3}$. We measure the time between
iterations $1$ and $32$, and average for the overhead per step. Every such
estimate is repeated over $10$ random seeds to obtain mean and error bars as
reported in \Cref{cockpit::fig:benchmark-instruments}.

Note that this protocol does \textit{not} include initial overhead for setting
up data loading and also does \textit{not} include the time for evaluating
train/test loss on a larger dataset, which is usually done by practitioners.
Hence, we even expect the shown overheads to be smaller in a conventional
training loop which includes the above steps.

\subsubsection{Individual Overhead on GPU Versus CPU}

\Cref{cockpit::fig:app_benchmark_instruments_cuda} and
\Cref{cockpit::fig:app_benchmark_instruments_cpu} show the individual overhead
for four different \deepobs problems on GPU and CPU, respectively. The left part
of \Cref{cockpit::fig:app_benchmark_instruments_cuda} (c) corresponds to
\Cref{cockpit::fig:benchmark-instruments}. Right panels show the expensive
quantities, which we omitted in the main text as they were expected to be
expensive due to their computational work (\inlinecode{HessMaxEV}) or
bottlenecks in the implementation (\inlinecode{GradHist2d}, see
\Cref{cockpit::app:histograms} for details). We see that they are in many cases
equally or more expensive than computing all other instruments. Another expected
feature of the GPU-to-CPU comparison is that parallelism on the CPU is
significantly less pronounced. Hence, we observe an increased overhead for all
quantities that contain non-linear transformations and contractions of the
high-dimensional individual gradients, or require additional backpropagations
(curvature).

\input{figures/cockpit/01_benchmark/app_gpu_individual_overhead}

\input{figures/cockpit/01_benchmark/app_cpu_individual_overhead}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End:
