\subsubsection{Procedure}

We use the checkpoints from \Cref{vivit::sec:ggn_vs_hessian}. For every
checkpoint, we compute the top-$C$ eigenvectors of the mini-batch \ggn{}
($N=128$) using an iterative matrix-free method. We also compute the mini-batch
gradient. The first- and second-order directional derivatives of the resulting
quadratic model (\Cref{vivit::eq:quadratic_model}) are given by
\Cref{vivit::eq:gammas-lambdas}.

We use these directional derivatives $\{\gamma_{n,k}\}_{n=1, k=1}^{N, C}$,
$\{\lambda_{n,k}\}_{n=1, k=1}^{N, C}$ to compute signal-to-noise ratios (SNRs)
along the top-$C$ eigenvectors. The curvature SNR along direction $\ve_k$ is
given by the squared sample mean divided by the empirical variance of the
samples $\{\lambda_{n,k}\}_{n=1}^{N}$, \ie
\begin{equation*}
  \text{SNR} = \frac{
    \lambda_k^2
  }{
    \nicefrac{1}{N-1} \sum_{n=1}^N (\lambda_{n,k} - \lambda_k)^2
  }
  \quad \text{where} \quad
  \lambda_k = \frac{1}{N} \sum_{n=1}^N \lambda_{n,k} \, .
\end{equation*}
(and similarly for $\{\gamma_{n,k}\}_{n=1}^{N}$).

\subsubsection{Results}

\Cref{vivit::fig:directional_derivatives_1,vivit::fig:directional_derivatives_2}
show the results.
%
These plots show the SNRs in $C$ distinct colors that generated from linear
interpolation in the RGB color space from black
(\tikz\draw[white,fill={black},line width=0mm] (0,0) circle (.8ex);) to light
red (\tikz\draw[white,fill={light_red},line width=0mm] (0,0) circle (.8ex);). At
each checkpoint, the colors are assigned based on the \textit{order} of the
respective directional curvature $\lambda_k$: the SNR that belongs to the
direction with the smallest curvature is shown in black and the SNR that belongs
to the direction with the largest curvature is shown in light red. The color
thus encodes only the order of the top-$C$ directional curvatures --
\textit{not} their magnitude. We use this color encoding to reveal potential
correlations between SNR and curvature.

We find that the gradient SNR along the top-$C$ eigenvectors is consistently
small (in comparison to the curvature SNR) and remains roughly on the same level
during the optimization. The curvature signal decreases as training proceeds.
The SNRs along the top-C eigendirections do not appear to show any significant
correlation with the corresponding curvatures. Only for the \cifarhun test
problems we can suspect a correlation between strong curvature and small
curvature SNR.

\input{figures/vivit/directional_derivatives}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End:
