% ===================================================================
% DISCLAIMER
% ===================================================================
{
  \thispagestyle{empty}
  \raggedright\null\vfill
  \paragraph{Disclaimer:} this thesis uses Federico Marotta's \texttt{kaobook}
  template based on Ken Arroyo Ohori's doctoral thesis. I am grateful to Frank
  Schneider who added many stylistic features and would like to thank him,
  Katharina Ott, Frederik Kunstner, Julia Grosse, Jonathan Wenger, Matthias
  Werner, Marvin Pf\"ortner, Runa Eschenhagen, and Nathanel Bosch for their help
  to improve the text.
  %
}
\cleardoublepage

% ===================================================================
% ACKNOWLEDGMENTS
% ===================================================================

\chapter{Acknowledgments}

% PROFESSIONAL PART

% Advisor
I want to express my deepest gratitude to Philipp Hennig for his guidance that
allowed me to seamlessly change fields for the \PhD I am inspired by your
passion for research, and grateful for your advice and everything I have learned
from you. Thank you for allowing me to travel the world, for spontaneously
stepping in on poster presentations, and for taking care of rebellious coffee
machines.

% Committee
Also, I am grateful to Asja Fischer and Hendrik Lensch who invested their time
and effort into reviewing this thesis, as well as Zeynep Akata and Georg Martius
for serving on my thesis advisory committee and forming my examination
committee.

% Administration
I want to thank
Franz{\tikzexternaldisable\TikzLowercaseSlightlySmilingFaceI{}}ska Weiler
(Methods of Machine Learning (MoML) group), Leila Masri \& Sara Sorce
(International Max Planck Research School for Intelligent Systems) for their
effective administrative support.

% Collaborators
This work was made possible by great collaborations. I want to thank Stefan
Harmeling, Frederik Kunstner, Frank Schneider, and Lukas Tatzel for these
experiences and their patience with me. I very much appreciated our regular
coffee breaks, coding sessions, ``grammar police'' discussions, and ``random
questions of today''.
% Lab members
This also applies more broadly to the entire MoML group that provided a
productive and friendly work environment. I consider myself lucky to have met
all of you
% : Motonobu Kanagawa, Filip Tronarp, Maren Mahsereci, Simon
% Bartels, Hans Kersting, Alexandra Gessner, Filip de Roos, Matthias Werner,
% Susanne Zabel, Agustinus Kristiadi, Jonathan Wenger, Nicholas Kr\"amer, Thomas
% Gl\"a\ss{}le, Kerstin Rau, Katharina Ott, Nathanael Bosch, Julia Grosse, Marius
% Hobbhahn, Emilia Magnani, Jonathan Schmidt, Andres Fernandez Rodriguez, Runa
% Eschenhagen, Marvin Pf\"ortner, Nina Effenberger, Tim Fischer, Ludwig Bald,
% Aaron Bahde, Hanna Dettki, and Elisabeth Knigge
(apologies for the regular mess
in the office kitchen).
% Students
I also enjoyed supervising Christian Meier, Paul Fischer, Shrisha Bharadwaj, Tim
Sch\"afer, Jessica Bader, and Elisabeth Knigge. Thank you for contributing to
this work; hopefully you learned as much as I did.

% PRIVATE PART
% Uni + MPI
During my time in T\"ubingen, I was fortunate enough to be surrounded by amazing
people. I want to express my gratitude to Jack Brady, Nicol\`o Zottino, Ori
Press, Emilia Magnani, Laura Iacovissi, Georg Hafner, Massimiliano Mancini, Ilya
Petrov, Faezeh Zakeri, Wafaa Mohammed, Martina Contisciani, Alessandro Lonardi,
Nicol\`o Ruggeri, Diego Theuerkauf, Ahmad Gazar, and everyone I forgot to list
here. You made this a really great time.

% Friends
Thank you to Jan Koschke, Marcel Bauer, and Jens Hollmann for being great
friends, and your support during---and long before---my studies.
Please do not ever call me Doctor D.\,(you know what I mean\dots).

% Family
I am grateful to my family, especially my parents, my sister, and my
grandparents (happy birthday, Oma!). They sparked my curiosity about science,
trusted in me, and provided me with everything I could hope for throughout the
years.
\begin{otherlanguage}{portuguese}
  Eu tamb\'em mando um abra\c{c}o forte \`{a} minha segunda fam\'ilia
  brasileira. Dona C\'icera, Senhor Jo\~{a}o Bosco, Gabriela, fico muito feliz
  por ter voc\^{e}s na minha vida.
\end{otherlanguage}

% Dani
Finally, I cannot thank Daniela Leite enough for her support during this
journey, and for being such a wonderful person. Thank you for showing me the
world from perspectives I would never have known otherwise. I hope the future
holds many more adventures and moments we share.

Thank you!

\begin{flushright}
  \textit{Felix Dangel}\\
  T\"ubingen, August 31, 2022
\end{flushright}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End:
